### 2021-02-25 code/draft.Rmd
updated draft.Rmd with a little more

---
title: "draft"
author: "Jincheng Wang"
date: "7/2/2020"
output: html_document
---

## Preparation
1. Setup
```{r setup, message=FALSE, warning=FALSE}
# Loading library here
library(knitr)
library(tidyverse)
library(readxl)
library(ggrepel)
# library(ggpubr)

options(max.print="75")
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE, prompt=FALSE, tidy = TRUE, comment = NA, message = FALSE)
opts_knit$set(width=75)

# set path
opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
setwd(rprojroot::find_rstudio_root_file())
## path to data record on dropbox
path_rec <- "~/Dropbox/IFNH CSR/ExperimentRecords/"
list.files(path_rec)

# number of cores/thread to use
nc <- switch(Sys.info()[['sysname']], Windows = 1, Darwin = 4, Linux = 12)

source("~/Dropbox/42-JCHWANG-RUTGERS/Projects-Rutgers/Src/utils.R")

sessionInfo()
```

## Body weight
### Overall, plotting all data
```{r}
BW <- read_excel(paste0(path_rec, "MasterMetadata-CS-SW-Oct2019-20210224-HS.xlsx"), sheet = "BodyWeightRecords", range = "A1:T2009") %>% filter(!is.na(Weight_g))
BW$BirthMode = factor(BW$BirthMode, levels = c("VF", "CSR", "CS"))

# BW <- BW %>% filter(!FamilyID %in% c("VF1", "CSR1") )
current_days <- max(BW$TimePoint_Days)

BW_sum <- BW %>% group_by(TimePoint_Days, BirthMode, Sex) %>% summarise(N = n(), Mean = mean(Weight_g), SE = sd(Weight_g)/sqrt(N))

y_max <- ceiling(max(BW$Weight_g)/4)*4
ggplot(BW, aes(x = TimePoint_Days, y = Weight_g, color = BirthMode)) +
    geom_point(size = 0.5, alpha = 0.5) +
    geom_path(aes(group = MouseNo), color = "grey", size = 0.5, alpha = 0.5) +
    geom_smooth(aes(fill = BirthMode), alpha = 0.3) +
    geom_text_repel(data = BW_sum, aes(x = TimePoint_Days, y = Mean, label = N), size = 4, show.legend = F) +
    facet_grid(.~Sex) +
    scale_x_continuous(breaks = c(seq(7, current_days, 7)), labels = c(seq(7, current_days, 7))/7) + 
    scale_y_continuous(breaks = seq(2, y_max, 4)) +
    scale_color_manual(values = c("#377EB8", "#4DAF4A", "#E41A1C")) +
    scale_fill_manual(values = c("#377EB8", "#4DAF4A", "#E41A1C")) +
    labs(x = "Sampling_Week", y = "Body Weight(g)", title = "C-Section Restoration Project - Body Weight Plots-AllData") +
    theme_jw() + theme(panel.grid.minor = element_blank())
```

### Summary table
```{r}
BW_sum2 <- BW %>% group_by(TimePoint_Weeks, BirthMode, Sex) %>% summarise(N = n(), N_batch = unique(Batch) %>% length, N_family = unique(FamilyNo) %>% length(), Mean = mean(Weight_g), SE = sd(Weight_g)/sqrt(N))

BW_sum2 %>% mutate(Count = paste0(N, "(", N_family, ")")) %>% select(BirthMode, Sex, TimePoint_Weeks, Count) %>% spread(key = TimePoint_Weeks, value = Count) %>% View
```

### Detailed plot
```{r}
date_range <- c(2, BW_sum2$TimePoint_Weeks[BW_sum2$N_family>=5] %>% max)

BW_dat <- BW %>% dplyr::filter(TimePoint_Weeks<=date_range[2], TimePoint_Weeks>=date_range[1])

BW_dat_stat_sum <- BW_dat %>% group_by(TimePoint_Weeks, BirthMode, Sex) %>% summarise(Mean_cl = Hmisc::smean.cl.normal(x = Weight_g) %>% paste(collapse = "-"), N = n()) %>% separate(Mean_cl, into = c("Mean", "Lwr", "Uwr"), sep = "-", convert = T) %>% ungroup

BW_dat_test_sum <- split(BW_dat, f = list(BW_dat$TimePoint_Weeks, BW_dat$Sex)) %>% lapply(., function(x){(aov(Weight_g ~ BirthMode, data = x) %>% agricolae::duncan.test(trt = "BirthMode", console = T))$groups %>% rownames_to_column(var = "BirthMode")})

BW_dat_test_sum2 <- lapply(seq_along(BW_dat_test_sum), function(x){BW_dat_test_sum[[x]] %>% mutate(Weeks_sex = names(BW_dat_test_sum)[[x]]) %>% separate(Weeks_sex, into = c("TimePoint_Weeks", "Sex"), sep = "\\.")}) %>% do.call("rbind",.) %>% mutate(TimePoint_Weeks = as.numeric(TimePoint_Weeks))

ggplot(BW_dat, aes(x = TimePoint_Weeks, y = Weight_g, color = BirthMode)) +
    geom_point(data = BW, size = 0.5, alpha = 0.5) +
    geom_path(data = BW, aes(group = MouseNo), color = "grey", size = 0.5, alpha = 0.5) +
    stat_summary(data = BW %>% filter(TimePoint_Weeks>=10), geom = "point", fun = "mean", size = 1) +
    stat_summary(data = BW %>% filter(TimePoint_Weeks>=10), geom = "line", fun = "mean", linetype = "dashed") +
    geom_ribbon(data = BW_dat_stat_sum, aes(x = TimePoint_Weeks, y = Mean, ymin = Lwr, ymax = Uwr, fill = BirthMode), alpha = 0.3) +
    geom_point(data = BW_dat_stat_sum, aes(x = TimePoint_Weeks, y = Mean), size = 1.5) +
    geom_line(data = BW_dat_stat_sum, aes(x = TimePoint_Weeks, y = Mean)) +
    geom_label_repel(data = BW_dat_test_sum2, aes(label = groups), size = 6/.pt, label.padding = 0.1) +
    facet_grid(.~Sex) +
    scale_x_continuous(breaks = c(seq(2, 18, 1))) + 
    scale_color_manual(values = c("#377EB8", "#4DAF4A", "#E41A1C")) +
    scale_fill_manual(values = c("#377EB8", "#4DAF4A", "#E41A1C")) +
    labs(x = "Sampling_Week", y = "Body Weight(g)", title = "C-Section Restoration Project - Body Weight", subtitle = str_wrap("Mean and 95% CI were plotted; ANOVA test with duncan's multiple range tests were performed at each week ", width = 120)) +
    theme_jw() + theme(panel.grid.minor = element_blank(), aspect.ratio = 0.8, plot.title = element_text(hjust = 0))
set_panel_size(file = "results/figs/exploratory/BW.pdf", width = unit(4, "in"), height = unit(3.2, "in"), useDingbats = F)
```

## Body Weight gain relative to week 3
```{r}
BW_f <- BW %>% filter(ExperimentSet=="Set1-BW", TimePoint_Weeks>=3) %>% arrange(TimePoint_Weeks) %>% group_by(MouseID) %>% mutate(Weight_gain = Weight_g - Weight_g[1])

date_range <- c(4, 16)

BW_f_dat <- BW_f %>% filter(TimePoint_Weeks<=date_range[2], TimePoint_Weeks>=date_range[1])

BW_f_dat %>% group_by(TimePoint_Weeks, BirthMode, Sex) %>% summarise(N = n()) %>% pivot_wider(names_from = c(BirthMode, Sex), values_from = N)

BW_f_dat_stat_sum <- BW_f_dat %>% group_by(TimePoint_Weeks, BirthMode, Sex) %>% summarise(Mean_cl = Hmisc::smean.cl.normal(x = Weight_gain) %>% paste(collapse = "-"), N = n()) %>% separate(Mean_cl, into = c("Mean", "Lwr", "Uwr"), sep = "-", convert = T) %>% ungroup

BW_f_dat_test <- BW_f_dat %>% group_by(TimePoint_Weeks, Sex) %>% t_test(Weight_gain ~ BirthMode, p.adjust.method = "none") %>% ungroup() %>% mutate(p.adj = p.adjust(p, method = "fdr"))

BW_f_dat_test_tbl1 <- BW_f_dat_test %>% pivot_wider(id_cols = c(Sex, group1, group2), names_from = TimePoint_Weeks, values_from = p.adj)

BW_f_dat_test_sum <- split(BW_f_dat, f = list(BW_f_dat$TimePoint_Weeks, BW_f_dat$Sex)) %>% lapply(., function(x){(aov(Weight_gain ~ BirthMode, data = x) %>% agricolae::duncan.test(trt = "BirthMode", console = T))$groups %>% rownames_to_column(var = "BirthMode")})

BW_f_dat_test_sum2 <- lapply(seq_along(BW_f_dat_test_sum), function(x){BW_f_dat_test_sum[[x]] %>% mutate(Weeks_sex = names(BW_f_dat_test_sum)[[x]]) %>% separate(Weeks_sex, into = c("TimePoint_Weeks", "Sex"), sep = "\\.")}) %>% do.call("rbind",.) %>% mutate(TimePoint_Weeks = as.numeric(TimePoint_Weeks))

ggplot(BW_f_dat, aes(x = TimePoint_Weeks, y = Weight_gain, color = BirthMode)) +
    geom_point(size = 0.5, alpha = 0.5) +
    geom_path(aes(group = MouseNo), color = "grey", size = 0.5, alpha = 0.5) +
    stat_summary(geom = "point", fun = "mean", size = 1) +
    stat_summary(geom = "line", fun = "mean", linetype = "dashed") +
    geom_ribbon(data = BW_f_dat_stat_sum, aes(x = TimePoint_Weeks, y = Mean, ymin = Lwr, ymax = Uwr, fill = BirthMode), alpha = 0.2, color = NA) +
    geom_point(data = BW_f_dat_stat_sum, aes(x = TimePoint_Weeks, y = Mean), size = 1.5) +
    geom_line(data = BW_f_dat_stat_sum, aes(x = TimePoint_Weeks, y = Mean)) +
    geom_label_repel(data = BW_f_dat_test_sum2, aes(label = groups), size = 6/.pt, label.padding = 0.1) +
    facet_grid(.~Sex) +
    scale_x_continuous(breaks = c(seq(2, 18, 1))) + 
    scale_color_manual(values = c("#377EB8", "#4DAF4A", "#E41A1C")) +
    scale_fill_manual(values = c("#377EB8", "#4DAF4A", "#E41A1C")) +
    labs(x = "Sampling_Week", y = "Body Weight(g)", title = "C-Section Restoration Project - Body Weight", subtitle = str_wrap("Mean and 95% CI were plotted starting from week 4; ANOVA test with duncan's multiple range tests were performed at each week between week 4 and 15.", width = 120)) +
    theme_jw() + theme(panel.grid.minor = element_blank(), aspect.ratio = 0.8, plot.title = element_text(hjust = 0))

```

## DEXA
```{r}
MM <- read_excel(paste0(path_rec, "MasterMetadata-CS-SW-Oct2019-20200701-JW.xlsx"), sheet = "MouseMetadata")

MM_dexa <- MM %>% filter(ExperimentSet=="Set1-BW", ReasonTermination %in% c("DEXA", "Week18Termination")) %>% mutate(TimePoint_Weeks = ifelse(ReasonTermination=="DEXA", 6, 18))

MM_dexa %>% group_by(TimePoint_Weeks, BirthMode, Sex) %>% summarise(N_family = length(unique(FamilyID)), N = length(unique(MouseID))) %>% mutate(Count = paste0(N, " (", N_family, ")")) %>% select(BirthMode, Sex, TimePoint_Weeks, Count) %>% spread(key = TimePoint_Weeks, value = Count) %>% View

spread(key = Sex, value = N)
```

## Behavior
```{r}
Beh <- read_excel(paste0(path_rec, "BehaviorAnimalMetadata-20190124-JW.xlsx"))

Beh %>% filter(Batch != "NA") %>% group_by(BirthMode, Sex) %>% summarise(N_family = length(unique(FamilyID)), N = length(unique(MouseNo))) %>%  mutate(Count = paste0(N, " (", N_family, ")")) 
```

## Feces

```{r}
MM <- read_excel(paste0(path_rec, "MasterMetadata-CS-SW-Oct2019-20200701-JW.xlsx"), sheet = "SampleMetadata")
MM_fec <- MM %>% filter(BodySite=="Feces")
MM_fec$TimePointInWeeks[MM_fec$TimePointInWeeks %in% c(11, 14)] = "11/14"

MM_fec %>% group_by(TimePointInWeeks, BirthMode, Sex) %>% summarise(N_family = length(unique(FamilyID)), N = length(unique(MouseID))) %>% mutate(Count = paste0(N, " (", N_family, ")")) %>% select(BirthMode, Sex, TimePointInWeeks, Count) %>% spread(key = TimePointInWeeks, value = Count) %>% View
```

## Sac
```{r}
MM <- read_excel(paste0(path_rec, "MasterMetadata-CS-SW-Oct2019-20200701-JW.xlsx"), sheet = "SampleMetadata")
Sac <- MM %>% filter(!SampleType %in% c("Feces", "Empty"))
Sac$BirthMode[Sac$BirthMode=="VD"] <- "VF"

Sac %>% group_by(TimePointInWeeks, BirthMode, Sex) %>% summarise(N_family = length(unique(FamilyID)), N = length(unique(MouseID))) %>% mutate(Count = paste0(N, " (", N_family, ")")) %>% select(BirthMode, Sex, TimePointInWeeks, Count) %>% spread(key = TimePointInWeeks, value = Count) %>% View

```